---
title: "Learning tidyverse"
author: "Duy Phan"
date: "July 20, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#install.packages("stargazer")
#install.packages("tidyverse")
library(tidyverse)
library(stargazer)
```
If we need to be explicit about where a function (or dataset) comes from, we'll use the special form package::function(). For example, ggplot2::ggplot() tells you explicitly that we're using the ggplot() function from the ggplot2 package.

Check some of the default dataset
```{r}
mpg = ggplot2::mpg
mpg
head(attitude)
```
Creating a ggplot

To plot mpg, run this code to put displ on the x-axis and hwy on the y-axis:

```{r}
ggplot(data = mpg)
#Line graph
ggplot(data = mpg)+
  geom_line(mapping = aes(x=displ, y=hwy))
#Bar graph
ggplot(data = mpg)+
  geom_bar(mapping = aes(x=class),fill = "purple")

#Point graph
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))

```
The plot shows a negative relationship between engine size (displ) and fuel efficiency (hwy). In other words, cars with big engines use more fuel. Does this confirm or refute your hypothesis about fuel efficiency and engine size?

With ggplot2, you begin a plot with the function ggplot(). ggplot() creates a coordinate system that you can add layers to. The first argument of ggplot() is the dataset to use in the graph. So ggplot(data = mpg) creates an empty graph, but it's not very interesting so I'm not going to show it here.

You complete your graph by adding one or more layers to ggplot(). The function geom_point() adds a layer of points to your plot, which creates a scatterplot. ggplot2 comes with many geom functions that each add a different type of layer to a plot. You'll learn a whole bunch of them throughout this chapter.

Each geom function in ggplot2 takes a mapping argument. This defines how variables in your dataset are mapped to visual properties. The mapping argument is always paired with aes(), and the x and y arguments of aes() specify which variables to map to the x and y axes. ggplot2 looks for the mapped variable in the data argument, in this case, mpg.

#Aesthetic mappings
Notice that there are some cars hat have bigger engine but still maintaning fuel efficiency. 
=>One way to test this hypothesis is to look at the class value for each car. A theory that these are hybrid cars.
You can add a third variable, like class, to a two dimensional scatterplot by mapping it to an aesthetic. An aesthetic is a visual property of the objects in your plot. Aesthetics include things like the size, the shape, or the color of your points. 

You can convey information about your data by mapping the aesthetics in your plot to the variables in your dataset. For example, you can map the colors of your points to the class variable to reveal the class of each car.
```{r}
#Class by color
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = class))

#Class by size
#We get a warning here, because mapping an unordered variable (class) to an ordered aesthetic (size) is not a good idea.
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, size = class))

#Or we could have mapped class to the alpha aesthetic, which controls the transparency of the points, or the shape of the points.
# Alpha
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, alpha = class))

#Shape
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, shape = class))
# What happened to the SUVs? ggplot2 will only use six shapes at a time. By default, additional groups will go unplotted when you use the shape aesthetic.
```
You can also set the aesthetic properties of your geom manually. For example, we can make all of the points in our plot blue:
#use quatation mark for the color so that R will not understand "blue" as variable
```{r}
ggplot(data = mpg) + 
  geom_point (mapping = aes(x = displ, y = hwy, color = "pink"))
```
#What's gone wrong with this code? Why are the points not blue?
The points are not blue because the "blue" is being interpreted as a vector (c("blue")) to map to an aesthetic, just like hwy or displ. To manually override a colour, the mapping could be placed outside the aes.
```{r}
ggplot (data = mpg) +
  geom_point(mapping = aes(x=displ, y=hwy), color = "blue")
```
```{r}
#data description
?mpg
```
#Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables?
=>Spectrum of color
```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ, y = hwy, color = cty))
```

#Using mutiple aesthetic
```{r}
#shape by class and color red
ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ, y=hwy, shape = class), color ="red")
#color by city miles per gallon and using specific shape (fill fuction is not woking => maybe conflict with color)
ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ, y=hwy, color = cty), shape =7,fill="red")
#color by number of cylinders, size by city miles per gallon, shape by drive types, general color is red.
ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ, y=hwy, color = cyl, alpha = cty, shape = drv), color ="red")

```
What does the stroke aesthetic do? What shapes does it work with? (Hint: use ?geom_point)
=> Stroke only work with numeric variable. It control how big is the data point is (quite similar with alpha)
```{r}
#README
?geom_point
#README
vignette("ggplot2-specs")


#serious Example from the readme
p <- ggplot(mtcars, aes(mpg, wt, shape = factor(cyl)))
p + geom_point(aes(colour = factor(cyl)), size = 4) +
  geom_point(colour = "grey90", size = 1.5)
p + geom_point(colour = "black", size = 4.5) +
  geom_point(colour = "pink", size = 4) +
  geom_point(aes(shape = factor(cyl)))

#Stroke
ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ, y=hwy, stroke = cyl))

ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ, y=hwy, color=class, alpha = cyl), stroke= 3)
```
What happens if you map an aesthetic to something other than a variable name, like aes(colour = displ < 5)?
=> Highlight the arguement using color
```{r}
#Coloring cars that have engine bigger than 5 litres
ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ, y=hwy, shape = drv, color = displ>5))

#Finding cars that have big engine and still have high highway per gallon => By coloring everything else?
ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ, y=hwy, shape = drv, color = displ<5|hwy<22))
```
#Facets

One way to add additional variables is with aesthetics. Another way, particularly useful for categorical variables, is to split your plot into facets, subplots that each display one subset of the data.

To facet your plot by a single variable, use facet_wrap(). The first argument of facet_wrap() should be a formula, which you create with ~ followed by a variable name (here "formula" is the name of a data structure in R, not a synonym for "equation"). The variable that you pass to facet_wrap() should be discrete.
```{r}
ggplot(data=mpg)+
  geom_point(mapping = aes(x=displ, y=hwy))+
  facet_wrap(~class, nrow =1)

ggplot(data=mpg)+
  geom_point(mapping = aes(x=displ, y=hwy))+
  facet_wrap(~class, nrow =3)
```
To facet your plot on the combination of two variables, add facet_grid() to your plot call. The first argument of facet_grid() is also a formula. This time the formula should contain two variable names separated by a ~.
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid(drv ~ cyl)
```
If you prefer to not facet in the rows or columns dimension, use a . instead of a variable name, e.g. + facet_grid(. ~ cyl).
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid(.~cyl)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid(cyl~.)
```
#Practicing facet grid
1. What happens if you facet on a continuous variable? => Generate many grid
```{r}
ggplot(data =mpg)+
  geom_point(mapping=aes(x=displ,y=hwy))+
  facet_grid(~cty)

```

2. What do the empty cells in plot with facet_grid(drv ~ cyl) mean? How do they relate to this plot? => There are no data fall into those bracklets
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = drv, y = cyl))

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid(drv~cyl)
```

3.What plots does the following code make? What does . do?
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)
```

4.Take the first faceted plot in this section:
What are the advantages to using faceting instead of the colour aesthetic? What are the disadvantages? How might the balance change if you had a larger dataset?
=> Well. Facets show clearer trend for each classes but make it a bit difficult to see the over all trend of the whole sample.
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```

5.Read ?facet_wrap. What does nrow do? What does ncol do? What other options control the layout of the individual panels? Why doesn't facet_grid() have nrow and ncol arguments?
=> Because the number of row and collum in facet_grid depend on the variables used in arguement
```{r}
?facet_wrap

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(cyl~ class, nrow = 4, ncol=5)
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid(cyl~ class)

```
6.When using facet_grid() you should usually put the variable with more unique levels in the columns. Why?
=> It will look more clean and easier to see trend
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid(cyl~ class)
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid(class~cyl)
```
#Geometric objects
A geom is the geometrical object that a plot uses to represent data. People often describe plots by the type of geom that the plot uses. For example, bar charts use bar geoms, line charts use line geoms, boxplots use boxplot geoms, and so on. Scatterplots break the trend; they use the point geom. As we see above, you can use different geoms to plot the same data. The plot on the left uses the point geom, and the plot on the right uses the smooth geom, a smooth line fitted to the data.
```{r}
# left
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))

# right
ggplot(data = mpg) + 
  geom_smooth(mapping = aes(x = displ, y = hwy))
```
Every geom function in ggplot2 takes a mapping argument. However, not every aesthetic works with every geom. You could set the shape of a point, but you couldn't set the "shape" of a line. On the other hand, you could set the linetype of a line. geom_smooth() will draw a different line, with a different linetype, for each unique value of the variable that you map to linetype.
```{r}
ggplot(data = mpg) + 
  geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv))
```
If this sounds strange, we can make it more clear by overlaying the lines on top of the raw data and then coloring everything according to drv.
```{r}
ggplot(data = mpg) + 
  geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv))+
  geom_point(mapping = aes(x=displ, y=hwy, color=drv))
```
Notice that this plot contains two geoms in the same graph! If this makes you excited, buckle up. In the next section, we will learn how to place multiple geoms in the same plot.

#ggplot2 provides over 30 geoms, and extension packages provide even more (see https://www.ggplot2-exts.org for a sampling). The best way to get a comprehensive overview is the ggplot2 cheatsheet, which you can find at http://rstudio.com/cheatsheets. To learn more about any single geom, use help: ?geom_smooth.

Many geoms, like geom_smooth(), use a single geometric object to display multiple rows of data. For these geoms, you can set the group aesthetic to a categorical variable to draw multiple objects. ggplot2 will draw a separate object for each unique value of the grouping variable. In practice, ggplot2 will automatically group the data for these geoms whenever you map an aesthetic to a discrete variable (as in the linetype example). It is convenient to rely on this feature because the group aesthetic by itself does not add a legend or distinguishing features to the geoms.
```{r}
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy))
              
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy, group = drv))
    
ggplot(data = mpg) +
  geom_smooth(
    mapping = aes(x = displ, y = hwy, color = drv),
    show.legend = FALSE
  )
```
To display multiple geoms in the same plot, add multiple geom functions to ggplot():
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  geom_smooth(mapping = aes(x = displ, y = hwy))
```

```{r}
stargazer(mpg)

attitude

stargazer(attitude[1:4,], summary=FALSE, rownames=FALSE)
```

% Table created by stargazer v.5.2.2 by Marek Hlavac, Harvard University. E-mail: hlavac at fas.harvard.edu
% Date and time: Tue, Jul 24, 2018 - 3:19:54 PM
\begin{table}[!htbp] \centering 
  \caption{} 
  \label{} 
\begin{tabular}{@{\extracolsep{5pt}} ccccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
rating & complaints & privileges & learning & raises & critical & advance \\ 
\hline \\[-1.8ex] 
$43$ & $51$ & $30$ & $39$ & $61$ & $92$ & $45$ \\ 
$63$ & $64$ & $51$ & $54$ & $63$ & $73$ & $47$ \\ 
$71$ & $70$ & $68$ & $69$ & $76$ & $86$ & $48$ \\ 
$61$ & $63$ & $45$ & $47$ & $54$ & $84$ & $35$ \\ 
\hline \\[-1.8ex] 
\end{tabular} 
\end{table} 

